---
title: '"National Parks"'
author: "Margarita Alfaro"
date: "5/10/2022"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../output', warning = FALSE, message = FALSE)
```

# Introduction 

On August 25, 1916, President Woodrow Wilson signed the "Organic Act" creating the National Park Service, a federal bureau in the Department of the Interior responsible for maintaining national parks and monuments that were then managed by the department. The National Park System has since expanded to 423 units (often referred to as parks), more than 150 related areas, and numerous programs that assist in conserving the nation's natural and cultural heritage for the benefit of current and future generations.(national park gov)

# List of National Parks
There are 63 National Parks and 9 of them are in California which make it the state with more national parks followed by Alaska with 8.

1. Denali National Park, Alaska
2. Gates of the Arctic National Park, Alaska
3. Glacier Bay National Park, Alaska
4. Katmai National Park, Alaska
5. Kenai Fjords National Park, Alaska
6. Kobuk Valley National Park, Alaska
7. Lake Clark National Park, Alaska
8. Wrangell-St. Elias National Park, Alaska
9. National Park of American Samoa, American Samoa
10. Grand Canyon National Park, Arizona
11. Petrified Forest National Park, Arizona
12. Saguaro National Park, Arizona
13. Hot Springs National Park, Arkansas
13. Channel Islands National Park, California
15. Death Valley National Park, California and Nevada
16. Joshua Tree National Park, California
17. Kings Canyon National Park, California
18. Lassen Volcanic National Park, California
19. Pinnacles National Park, California
20. Redwood National Park, California
21. Sequoia National Park, California
22. Yosemite National Park, California
23. Black Canyon of the Gunnison National Park, Colorado
24. Great Sand Dunes National Park, Colorado
25. Mesa Verde National Park, Colorado
26. Rocky Mountain National Park, Colorado
27. Biscayne National Park, Florida
28. Everglades National Park, Florida
29. Dry Tortugas National Park, Florida
30. Haleakalā National Park, Hawaii
31. Hawai'i Volcanoes National Park, Hawaii
32. Yellowstone National Park, Idaho, Montana, and Wyoming
33. Indiana Dunes National Park, Indiana
34. Mammoth Cave National Park, Kentucky
35. Acadia National Park, Maine
36. Isle Royale National Park, Michigan
37. Voyageurs National Park, Minnesota
38. Gateway Arch National Park, Missouri and Illinois
39. Glacier National Park, Montana
40. Great Basin National Park, Nevada
41. White Sands National Park, New Mexico
42. Carlsbad Caverns National Park, New Mexico
43. Great Smoky Mountains National Park, North Carolina and Tennessee
44. Theodore Roosevelt National Park, North Dakota
45. Cuyahoga Valley National Park, Ohio
46. Crater Lake National Park, Oregon
47. Badlands National Park, South Dakota
48. Wind Cave National Park, South Dakota
49. Congaree National Park, South Carolina
50. Big Bend National Park, Texas
51. Guadalupe Mountains National Park, Texas
52. Arches National Park, Utah
53. Bryce Canyon National Park, Utah
54. Canyonlands National Park, Utah
55. Capitol Reef National Park, Utah
56. Zion National Park, Utah
57. Virgin Islands National Park, Virgin Islands
58. Shenandoah National Park, Virginia
59. Olympic National Park, Washington
60. Mount Rainier National Park, Washington
61. North Cascades National Park, Washington
62. New River Gorge National Park and Preserve, West Virginia
63. Grand Teton National Park, Wyoming


# Load Libraries
```{r}
library(tidyverse)
library(here)
library(gganimate)
library(ggplot2)
library(reactable)
library(plotrix)
library(cowsay)
```

# Load Data
```{r}
park_visits <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv")
state_pop <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/state_pop.csv")
```


# Plot 1 (State Population)
US have 50 states and California is the state with most of the population.

list of the 10 most populated states in US:

1. California (Population: 39,613,493)
2. Texas (Population: 29,730,311)
3. Florida (Population: 21,944,577)
4. New York (Population: 19,299,981)
5. Pennsylvania (Population: 12,804,123)
6. Illinois (Population: 12,569,321)
7. Ohio (Population: 11,714,618)
8. Georgia (Population: 10,830,007)

9.North Carolina (Population: 10,701,022)
10. Michigan (Population: 9,992,427)

Population in States that have National Parks from 1900 to 2015
```{r}
pop <- state_pop %>%
  select(year, state, pop) %>% #select variables
  mutate(year, state, pop) %>% #create new columns
  filter(year == "1900" | year == "1910" | year == "1920" | year == "1930" | year == "1940" | year == "1950" | year == "1960" | year == "1970" | year == "1980" | year == "1990" | year == "2000" | year == "2005" | year == "2010" | year == "2015") %>% #extract/filter specific years
  filter(state == "CA" | state == "AK" | state == "UT" | state == "CO" | state == "WA" | state == "FL" | state == "AZ" | state == "HI" | state == "NM" | state == "SD" | state == "TX" | state == "MT" | state == "WY" | state == "WV" | state == "VA" | state == "VI" | state == "SC" | state == "OR" | state == "OH" | state == "ND" | state == "NC" | state == "TN" | state == "NV" | state == "MO" | state == "IL" | state == "MN" | state == "MI" | state == "ME" | state == "KY" | state == "IN" | state == "ID" | state == "AR" | state == "AS") %>% #extract/filter specific state name
drop_na() #drops rows with NAs from a specific column
```

```{r, fig.width= 15, fig.height=8}
pop %>%
  ggplot(aes(x= state, y= pop))+
  geom_point(color="firebrick3", size=4) + #create the plot
   labs(x= "States with National Parks", y= "Population Growth", #labels our x and y-axis
       subtitle= "Increment in Population from 1900-2015 ", #creates and label the subtitle
       caption = "Source: 2019-09-17 TidyTuesday  Data from the dataisplural/data.world")+
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )+
  theme_light() + #add a theme to the plot
   theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  )+ #change out plot elements
  theme(axis.text.x = element_text(angle=90)) #put x-axis vertically

```

# Plot 2 (CA Population)
Pollulation in CA from 1900 to 2015 
```{r}
popCA <- state_pop %>%
  select(year, state, pop) %>% #select variables
  mutate(year, state, pop) %>% #select new columns
  filter(year == "1900" | year == "1910" | year == "1920" | year == "1930" | year == "1940" | year == "1950" | year == "1960" | year == "1970" | year == "1980" | year == "1990" | year == "2000" | year == "2005" | year == "2010" | year == "2015") %>% #select specific years
  filter(state == "CA") %>% #select specific state name "CA"
drop_na() #drops rows with NAs from a specific column
```

```{r}
popCA %>%
  ggplot(aes(x= year, y= pop))+ 
  geom_col()+ #create the plot
  geom_smooth()+ #adds a best fit line
  labs(x= "States", y= "Population Growth", #labels our x and y-axis
       title = "California", #creates our title
       subtitle= "Change in Population from 1900-2015 ", #creates and label the subtitle
       caption = "Source: 2019-09-17 TidyTuesday  Data from the dataisplural/data.world")+
  theme_classic() #add a theme
```


# Plot 3 (Table for CA Park Visits)
```{r}
park <- park_visits %>% 
  filter(state == "CA") %>% #select state name "CA"
  filter(year == "2016") %>% #select year "2016"
  drop_na() #drops rows with NAs from a specific column
```

# Reactable Table
Creates interactive data tables for R.
The following interactive table includes information about National Parks, National Historic Site, National Monument, National Recreation Area, and National Preserve located in California for the year 2016.

Interactive table for National Parks in CA for 2016
```{r}
reactable(park) #create the interactive table
```


# Plot 4 (National Parks in CA)
Visitors in 8/9 national parks in CA for 2016

1. Channel Islands National Park
2. Death Valley National Park
3. Joshua Tree National Park
4. Kings Canyon National Park
5. Lassen Volcanic National Park
6. Pinnacles National Park
7. Sequoia National Park
8. Yosemite National Park
9. ***(Redwood National Park) Not in Data


```{r}
CApark <- park_visits %>%
  filter(state == "CA") %>% #select specific state CA
  filter(year == "2016") %>% #select specific year
  filter(unit_type == "National Park") %>% #select national park only
  drop_na() #drops rows with NAs from a specific column
```

Data for the 3D pie graph
```{r}
CA <- c(1254688, 364807, 607479, 2505286, 215555, 1296283, 536068, 5028868) #vector containing the numeric values
labels <- c("Sequoia", "Channel Islands", "Kings Canyon", "Joshua Tree", "Pinnacles", "Death Valley", "Lassen Volcanic", "Yosemite") #add names to the 3D pie
```

```{r,fig.width= 6, fig.height=5.4}
pie3D(CA, labels = labels, #add the names
      explode = 0.1, #separate the pie into pieces
      main= "National Parks Visits in CA", #give tittle to the pie
      col=c("deepskyblue2", "goldenrod1", "pink1", "darkolivegreen3", "burlywood4", "azure3", "chocolate2", "cadetblue4"), #add colors to the pie
      labelpos=NULL)
```
# End
```{r}
say("END", by = "cow")
```
